---
title       : Unemployment Dataset Mapping
subtitle    : User-interactive color mapping within contiguous 48 US states   
author      : Gary W. Deinert
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Purpose:
### Interactive shiny app to color maps from the "unemp" dataset.   

### 4 Interactive Features:
+ User map title
+ Range slider (two values) to select unemployment range
+ State selector (selectizeInput)
+ Update button to make reactive changes    

### Calculations performed include:
+ Reconfigure db to include only selected states, counties within unemp % range
+ Match colors to appropriate unemployment figures
+ Enter the user's title 
+ Calculate states' average unemployment
+ Summarize states' total population

--- .class #id 

## R Code to generate the maps:

```{r eval=FALSE}
            plot_db$colorBuckets <- as.numeric(cut(plot_db$unemp, c(0, 3, 6, 9, 12, 15, 100)))
            plot_db$colorBuckets[plot_db$unemp < (pct_range[1])]=100
            plot_db$colorBuckets[plot_db$unemp > (pct_range[2])]=100            
            
            map("county", regions = states, col = colors[plot_db$colorBuckets], 
                fill = TRUE,resolution = 0,lty = 0, projection = "polyconic")
            title(map.title)
            leg.txt <- c("<3%", "3-6%", "6-9%", "9-12%", "12-15%", "15+%")
            legend("bottomright", leg.txt, horiz = FALSE, fill = colors)
            
            ## Add border around each county
            map("county", regions = states, col = "darkgrey", fill = FALSE, add = TRUE, 
                lty = 1, lwd = 0.2, projection = "polyconic")
```

---

### Example Results: (open "index.Rmd" to see original code)
```{r echo=FALSE, warning=FALSE}
library(shiny)
library(maps)
library(mapproj)
library(RColorBrewer)
library(plyr)
data(countyMapEnv)
data(county.fips)
data(stateMapEnv)
data(unemp)

working_db <- NULL
states.shown <- c("ohio","west virginia","pennsylvania")
map.title <- "Over 8% Unemployment in OH,WV,PA"
pct_range <- c(8,20)

unemp_county <- county.fips$polyname[match(unemp$fips,county.fips$fips)]      
unemp <- cbind(unemp,unemp_county)
unemp <- unemp[complete.cases(unemp),]
unemp$unemp_county <- as.character(unemp$unemp_county)
unemp$state <- sapply(strsplit(unemp$unemp_county,","),"[",1)
unemp$count <- 1

working_db <- unemp[unemp$state %in% states.shown,]

colors <- brewer.pal(6,"PuBuGn") 
                        
working_db$colorBuckets <- as.numeric(cut(working_db$unemp, c(0, 3, 6, 9, 12, 15, 100)))
working_db$colorBuckets[working_db$unemp < (pct_range[1])]=100
working_db$colorBuckets[working_db$unemp > (pct_range[2])]=100            
            
map("county", regions = states.shown, col = colors[working_db$colorBuckets], 
      fill = TRUE,resolution = 0,lty = 0, projection = "polyconic")
      title(map.title)
      leg.txt <- c("<3%", "3-6%", "6-9%", "9-12%", "12-15%", "15+%")
      legend("bottomright", leg.txt, horiz = FALSE, fill = colors)
            
## Add border around each county
map("county", regions = states.shown, col = "darkgrey", fill = FALSE, add = TRUE, 
      lty = 1, lwd = 0.2, projection = "polyconic")
            
avg_unemp <- aggregate(unemp ~ state, data=working_db,mean)
total <- aggregate(pop ~ state, data=working_db,sum)
counties <- aggregate(count ~ state, data=working_db,sum)
summary_df <- cbind(avg_unemp,total$pop,counties$count)
summary_df

```

---

## Lessons Learned from Project    

+ Shiny, shinyapps.io, slidify, html usage within shiny, UI reactivity
+ Panels, tabsets, user help, page layouts
+ Use of "browser()" to troubleshoot shiny programs (critical)
+ Reactivity and how to program it effectively
+ This is great stuff!
